---
title: "One approach to streamlining your analytical workflows using functions and for loops"
author: Sam Csik
date: April 14, 2023
theme: minty
format: html
---

## Introduction

Some text about the benefits of using functions & for loops

Let's say I want to look at ocean temperature data from three coastal sites within the Santa Barbara Coastal (SBC) LTER: Alegria Reef, Mohawk Reef, Carpenteria Reef. Data are collected by moored instrumentation (CTD & ADCP) at each site. Raw data are available for download on the EDI Data Portal:

  1. Alegria Reef: <https://portal.edirepository.org/nis/mapbrowse?packageid=knb-lter-sbc.2008.14>
  2. Mohawk Reef: <https://doi.org/10.6073/pasta/cbe43646b801bf6ee5231c301ea23f51>
  3. Carpenteria Reef: <https://portal.edirepository.org/nis/mapbrowse?packageid=knb-lter-sbc.2004.26>

## **Step 1:** Clean and plot *one* data set 

I (personally) find it absolutely critical to figure out how to clean/wrangle/tidy one data set on its own before I even begin thinking about writing a function. The data set I'm using contains *lots* of variables that I'm not interested in. Additionally....

#### **Load libraries**
```{r}
#| eval: true
#| echo: true
#| message: false
#| warning: false
library(tidyverse)
library(naniar)
```

#### **Import raw data** 

Either download from EDI directly: 
```{r}
#| eval: true
#| echo: true
#| message: false
#| warning: false
# ale <- read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=knb-lter-sbc.2008.14&entityid=15c25abf9bb72e2017301fa4e5b2e0d4")
# mko <- read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=knb-lter-sbc.2007.16&entityid=02629ecc08a536972dec021f662428aa")
# car <- read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=knb-lter-sbc.2004.26&entityid=1d7769e33145ba4f04aa0b0a3f7d4a76")
```

OR download files (they are large!) from Google Drive and save to `data/raw_data/sbc_lter_temps`
```{r}
#| eval: true
#| echo: true
#| message: false
#| warning: false
alegria <- read_csv(here::here("data", "raw_data", "alegria_mooring_ale_20210617.csv"))
mohawk <- read_csv(here::here("data", "raw_data", "mohawk_mooring_mko_20220330.csv"))
carpenteria <- read_csv(here::here("data", "raw_data", "carpinteria_mooring_car_20220330.csv"))
```

#### **Clean data** 
```{r}
#| eval: true
#| echo: true
#| message: false
#| warning: false
alegria_clean <- alegria |> 
  
  # keep only necessary columns
  select("year", "month", "day", "decimal_time", "Temp_top", "Temp_mid", "Temp_bot") |>
  
  # add column with site name
  mutate(site = rep("Alegria Reef")) |> 
  
  # create date time column
  unite(date, year, month, day, sep = "-", remove = FALSE) |> 
  mutate(time = chron::times(as.numeric(decimal_time))) |> 
  unite(date_time, date, time, sep = " ", remove = TRUE) |> 
  
  # coerce data types
  mutate(date_time = as.POSIXct(date_time, "%Y-%m-%d %H:%M:%S", tz = "GMT"),
         year = as.factor(year),
         month = as.factor(month),
         day = as.numeric(day)) |>
  
  # add month name
  mutate(month_name = as.factor(month.name[month])) |>
  
  # replace 9999s with NAs (will throw warning if var isn't present, but still execute)
  replace_with_na(replace = list(Temp_bot = 9999, Temp_top = 9999, Temp_mid = 9999)) |> 
  
  # select/reorder desired columns
  select("site", "date_time", "year", "month", "day", "month_name", "Temp_bot", "Temp_mid", "Temp_top")

```

#### plot data
```{r}
#| eval: true
#| echo: true
#| message: false
#| warning: false
alegria_plot <- alegria_clean |> 
  
  group_by(month_name) |> 
  
    ggplot(aes(x = Temp_bot, y = month_name, fill = after_stat(x))) +
    ggridges::geom_density_ridges_gradient(rel_min_height = 0.01, scale = 3) + 
    scale_x_continuous(breaks = c(9, 12, 15, 18, 21)) +
    scale_y_discrete(limits = rev(month.name)) + 
    scale_fill_gradientn(colors = c("#2C5374","#778798", "#ADD8E6", "#EF8080", "#8B3A3A"), name = "Temp. (°C)") +
    labs(x = "Bottom Temperature (°C)",
         title = "Bottom Temperatures at Alegria Reef, Santa Barbara, CA",
         subtitle = "Temperatures (°C) aggregated by month from 2005 - 2022") +
    ggridges::theme_ridges(font_size = 13, grid = TRUE) +
    theme(
      axis.title.y = element_blank()
    )

alegria_plot
```

## **Step 2:** Write functions to clean & plot your data

### **Step 2.1** Write a function to **clean** data sets

### **Step 2.2:** Write a function to **plot** data sets

### **Step 2.3:** Use your functions to **clean and plot** your data sets



