---
format: html
---

# **`r fontawesome::fa("person-walking", fill = "#5A5A5A", a11y = "sem")` Stage 1:** Clean and plot *one* data set 

Copying, pasting, and slightly updating code to clean and process multiple similarly-structured data sets is certainly the slowest and most prone-to-errors workflow. However, (in *my* honest opinion) it's almost always critical to being by processing one data set on it's own before attempting to write a function to do so.

Let's start by doing that here.

::: {.callout-note}
I'll be writing the following code in a script called, `no_functions_pipeline.R`, which I'll save to my repo's root directory.
:::

## **i.** Load libraries
```{r}
#| eval: true
#| echo: true
#| message: false
#| warning: false
library(tidyverse)
library(chron)
library(naniar)
library(ggridges)
```

## **ii.** Import raw data 

Either download from the [Environmental Data Portal](https://portal.edirepository.org/nis/home.jsp) (EDI) directly: 
```{r}
#| eval: false
#| echo: true
#| message: false
#| warning: false
# ale <- read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=knb-lter-sbc.2008.14&entityid=15c25abf9bb72e2017301fa4e5b2e0d4")
# mko <- read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=knb-lter-sbc.2007.16&entityid=02629ecc08a536972dec021f662428aa")
# car <- read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=knb-lter-sbc.2004.26&entityid=1d7769e33145ba4f04aa0b0a3f7d4a76")
```

Or read in the downloaded files from `data/raw_data/`
```{r}
#| eval: true
#| echo: true
#| message: false
#| warning: false
alegria <- read_csv(here::here("data", "raw_data", "alegria_mooring_ale_20210617.csv"))
mohawk <- read_csv(here::here("data", "raw_data", "mohawk_mooring_mko_20220330.csv"))
carpinteria <- read_csv(here::here("data", "raw_data", "carpinteria_mooring_car_20220330.csv"))
```

## **iii.** Clean data 

Below, we select only the necessary columns (there are *far* too many (87) in the raw data), add a column for site name (the only way to tell which site the data were collected from is by looking at the file name), formatting dates/times, and replacing missing value codes with *`NA`*.
```{r}
#| eval: true
#| echo: true
#| message: false
#| warning: false
alegria_clean <- alegria |> 
  
  # keep only necessary columns & filter for years 2005-2020
  select(year, month, day, decimal_time, Temp_top, Temp_mid, Temp_bot) |>
  filter(year %in% c(2005:2020)) |> 
  
  # add column with site name
  mutate(site = rep("Alegria Reef")) |> 
  
  # create date time column
  unite(col = date, year, month, day, sep = "-", remove = FALSE) |> 
  mutate(time = times(decimal_time)) |> 
  unite(col = date_time, date, time, sep = " ", remove = TRUE) |> 
  
  # coerce data types
  mutate(date_time = as.POSIXct(date_time, "%Y-%m-%d %H:%M:%S", tz = "GMT"),
         year = as.factor(year),
         month = as.factor(month),
         day = as.numeric(day)) |>
  
  # add month name (see: https://stackoverflow.com/questions/22058393/convert-a-numeric-month-to-a-month-abbreviation)
  mutate(month_name = as.factor(month.name[month])) |>
  
  # replace 9999s with NAs (will throw warning if var isn't present, but still execute)
  replace_with_na(replace = list(Temp_bot = 9999, Temp_top = 9999, Temp_mid = 9999)) |> 
  
  # select/reorder desired columns
  select(site, date_time, year, month, day, month_name, Temp_bot, Temp_mid, Temp_top)

```

## **iv.** Plot the data

Here, we create a ridge line plot (using the `{ggridges}` package) showing aggregate bottom temperatures (2005-2020), by month, at Alegria reef.
```{r}
#| eval: true
#| echo: true
#| message: false
#| warning: false
alegria_plot <- alegria_clean |> 
  
  # group by month ----
  group_by(month_name) |> 
  
  # plot ----
  ggplot(aes(x = Temp_bot, y = month_name, fill = after_stat(x))) +
  geom_density_ridges_gradient(rel_min_height = 0.01, scale = 3) + 
  scale_x_continuous(breaks = c(9, 12, 15, 18, 21)) +
  scale_y_discrete(limits = rev(month.name)) + 
  scale_fill_gradientn(colors = c("#2C5374","#778798", "#ADD8E6", "#EF8080", "#8B3A3A"), name = "Temp. (°C)") +
  labs(x = "Bottom Temperature (°C)",
       title = "Bottom Temperatures at Alegria Reef, Santa Barbara, CA",
       subtitle = "Temperatures (°C) aggregated by month from 2005 - 2020") +
  ggridges::theme_ridges(font_size = 13, grid = TRUE) +
  theme(
    axis.title.y = element_blank()
  )

alegria_plot
```

If we were to continue with this workflow (which is absolutely a valid way that gets the job done!), we would need to repeat the above code two more times (for both the `mohawk` and `carpinteria` data frames) -- this gets lengthy rather quickly, requires lots of copying/pasting, and is prone to errors (e.g. forgetting to update a data frame name, typos, etc.). If you'd like to check out the code for the `mohawk` and `carpinteria` data sets, unfold the code chunk below:

```{r}
#| eval: true
#| echo: true
#| code-fold: true
#| layout-ncol: 2
#| message: false
#| warning: false

#..........................Mohawk Reef...........................

# clean ----
mohawk_clean <- mohawk |> 
  select(year, month, day, decimal_time, Temp_top, Temp_mid, Temp_bot) |>
  filter(year %in% c(2005:2020)) |> 
  mutate(site = rep("Mohawk Reef")) |> 
  unite(col = date, year, month, day, sep = "-", remove = FALSE) |> 
  mutate(time = times(decimal_time)) |> 
  unite(col = date_time, date, time, sep = " ", remove = TRUE) |> 
  mutate(date_time = as.POSIXct(date_time, "%Y-%m-%d %H:%M:%S", tz = "GMT"),
         year = as.factor(year),
         month = as.factor(month),
         day = as.numeric(day)) |>
  mutate(month_name = as.factor(month.name[month])) |>
  replace_with_na(replace = list(Temp_bot = 9999, Temp_top = 9999, Temp_mid = 9999)) |> 
  select(site, date_time, year, month, day, month_name, Temp_bot, Temp_mid, Temp_top)

# plot ----
mohawk_plot <- mohawk_clean |> 
  group_by(month_name) |> 
  ggplot(aes(x = Temp_bot, y = month_name, fill = after_stat(x))) +
  geom_density_ridges_gradient(rel_min_height = 0.01, scale = 3) + 
  scale_x_continuous(breaks = c(9, 12, 15, 18, 21)) +
  scale_y_discrete(limits = rev(month.name)) + 
  scale_fill_gradientn(colors = c("#2C5374","#778798", "#ADD8E6", "#EF8080", "#8B3A3A"), name = "Temp. (°C)") +
  labs(x = "Bottom Temperature (°C)",
       title = "Bottom Temperatures at Mohawk Reef, Santa Barbara, CA",
       subtitle = "Temperatures (°C) aggregated by month from 2005 - 2020") +
  ggridges::theme_ridges(font_size = 13, grid = TRUE) +
  theme(
    axis.title.y = element_blank()
  )

mohawk_plot

#........................Carpinteria Reef........................

# clean ----
carpinteria_clean <- carpinteria |> 
  select(year, month, day, decimal_time, Temp_top, Temp_mid, Temp_bot) |>
  filter(year %in% c(2005:2020)) |> 
  mutate(site = rep("Carpinteria Reef")) |> 
  unite(col = date, year, month, day, sep = "-", remove = FALSE) |> 
  mutate(time = times(decimal_time)) |> 
  unite(col = date_time, date, time, sep = " ", remove = TRUE) |> 
  mutate(date_time = as.POSIXct(date_time, "%Y-%m-%d %H:%M:%S", tz = "GMT"),
         year = as.factor(year),
         month = as.factor(month),
         day = as.numeric(day)) |>
  mutate(month_name = as.factor(month.name[month])) |>
  replace_with_na(replace = list(Temp_bot = 9999, Temp_top = 9999, Temp_mid = 9999)) |> 
  select(site, date_time, year, month, day, month_name, Temp_bot, Temp_mid, Temp_top)

# plot ----
carpinteria_plot <- carpinteria_clean |> 
  group_by(month_name) |> 
  ggplot(aes(x = Temp_bot, y = month_name, fill = after_stat(x))) +
  geom_density_ridges_gradient(rel_min_height = 0.01, scale = 3) + 
  scale_x_continuous(breaks = c(9, 12, 15, 18, 21)) +
  scale_y_discrete(limits = rev(month.name)) + 
  scale_fill_gradientn(colors = c("#2C5374","#778798", "#ADD8E6", "#EF8080", "#8B3A3A"), name = "Temp. (°C)") +
  labs(x = "Bottom Temperature (°C)",
       title = "Bottom Temperatures at Carpinteria Reef, Santa Barbara, CA",
       subtitle = "Temperatures (°C) aggregated by month from 2005 - 2020") +
  ggridges::theme_ridges(font_size = 13, grid = TRUE) +
  theme(
    axis.title.y = element_blank()
  )

carpinteria_plot
```