---
format: html
---

# **`r fontawesome::fa("person-running", fill = "#5A5A5A", a11y = "sem")` Stage 2:** Write functions to clean & plot your data

Though there isn't anything inherently *wrong* with copying/pasting large chunks of code, it's a better practice to turn repeated code into functions. In Stage 2, we'll turn our cleaning and plotting pipelines into functions that can be called on any of our similarly-structured data sets.

::: {.callout-note}
## Where do I write/save my functions?
While there's no hard and fast rule, I tend to create a subdirectory within my project (e.g. named `/R`, `/utils`, etc.) to house all my function scripts. I prefer to create a separate `.R` script for each of my functions, and name each file the same as the function itself (e.g. if I'm writing a function called `do_fun_thing()`, I'd save it to a script called `do_fun_thing.R`).

You can then **source** your function files into whatever script (or `.rmd`/`.qmd` file) where you call that function (e.g. `source("utils/do_fun_thing.R")`. 

**Note:** You *cannot* `source()` a `.rmd` or `.qmd` file into another file/script, therefore it's important to save functions to a `.R` file.

Here, let's create the following:

-  a `/utils` folder to store our functions scripts
-  a `clean_ocean_temps.R` file (saved to `/utils`), where we'll write a function to clean our data
-  a `plot_ocean_temps.R` file (saved to `/utils`), where we'll write a function to plot our data
-  a `functions_pipeline.R` file (saved to the root directory), where we'll use our functions to clean and plot our data -- you can also read-in your data files here, just as we did in [Stage 1, part ii](http://localhost:7074/#ii.-import-raw-data)
:::

## **i.** Write a function to **clean** data sets

It's helpful to first identify which parts of the cleaning code need to be generalized/made "flexible" so that any of our three data frames can be passed to it for cleaning. For us, that's the **name of the raw data frame** and the **site name character string** that's repeated for the length of the added `site` column (see annotation notes below the rendered code):

```{r}
#| eval: false
#| echo: true
#| message: false
#| warning: false
alegria_clean <- alegria |> # <1>
  select(year, month, day, decimal_time, Temp_top, Temp_mid, Temp_bot) |>
  filter(year %in% c(2005:2020)) |> 
  mutate(site = rep("Alegria Reef")) |> # <2>
  unite(col = date, year, month, day, sep = "-", remove = FALSE) |> 
  mutate(time = times(decimal_time)) |> 
  unite(col = date_time, date, time, sep = " ", remove = TRUE) |> 
  mutate(date_time = as.POSIXct(date_time, "%Y-%m-%d %H:%M:%S", tz = "GMT"),
         year = as.factor(year),
         month = as.factor(month),
         day = as.numeric(day)) |>
  mutate(month_name = as.factor(month.name[month])) |>
  replace_with_na(replace = list(Temp_bot = 9999, Temp_top = 9999, Temp_mid = 9999)) |> 
  select(site, date_time, year, month, day, month_name, Temp_bot, Temp_mid, Temp_top)
```
1. the name of the raw data frame (here, `alegria`) needs to be generalized
2. the site name character string that's repeated for the length of the added `site` column (here, **"Alegria Reef"**) needs to be generalized

Now we can start to build out our function. We'll start by creating a super basic function, then build in more complexity. I encourage you to test out your function after each version to ensure that it works as you intend it to.

### **Version 1:** 

***The primary goal of `clean_ocean_temps()` v1 is to get a basic data cleaning function working.*** 

To start, let's create the skeleton of our function, which we'll call `clean_ocean_temps()`:

```{r}
#| eval: false
#| echo: true
#| message: false
#| warning: false
clean_ocean_temps <- function(){
  
}
```

Now, let's copy and paste our cleaning code for Alegria Reef data (from **Stage 1**, above) into the body of the function (i.e. within the curly brackets, `{}`).

```{r}
#| eval: false
#| echo: true
#| message: false
#| warning: false
clean_ocean_temps <- function(){
 
  alegria_clean <- alegria |> 
    select(year, month, day, decimal_time, Temp_top, Temp_mid, Temp_bot) |>
    filter(year %in% c(2005:2020)) |> 
    mutate(site = rep("Alegria Reef")) |> 
    unite(col = date, year, month, day, sep = "-", remove = FALSE) |> 
    mutate(time = times(decimal_time)) |> 
    unite(col = date_time, date, time, sep = " ", remove = TRUE) |> 
    mutate(date_time = as.POSIXct(date_time, "%Y-%m-%d %H:%M:%S", tz = "GMT"),
          year = as.factor(year),
          month = as.factor(month),
          day = as.numeric(day)) |>
    mutate(month_name = as.factor(month.name[month])) |>
    replace_with_na(replace = list(Temp_bot = 9999, Temp_top = 9999, Temp_mid = 9999)) |> 
    select(site, date_time, year, month, day, month_name, Temp_bot, Temp_mid, Temp_top)
  
}
```

Next, we want the ability to provide our function with any three of our data sets for processing. To do so, we'll make the following updates: 

```{r}
#| eval: false
#| echo: true
#| message: false
#| warning: false
clean_ocean_temps <- function(raw_data){ # <1>
  
  # clean data ----
  temps_clean <- raw_data |> # <2>
    select(year, month, day, decimal_time, Temp_top, Temp_mid, Temp_bot) |>
    filter(year %in% c(2005:2020)) |> 
    mutate(site = rep("_____")) |> # <4>
    unite(col = date, year, month, day, sep = "-", remove = FALSE) |> 
    mutate(time = times(decimal_time)) |> 
    unite(col = date_time, date, time, sep = " ", remove = TRUE) |> 
    mutate(date_time = as.POSIXct(date_time, "%Y-%m-%d %H:%M:%S", tz = "GMT"),
          year = as.factor(year),
          month = as.factor(month),
          day = as.numeric(day)) |>
    mutate(month_name = as.factor(month.name[month])) |>
    replace_with_na(replace = list(Temp_bot = 9999, Temp_top = 9999, Temp_mid = 9999)) |> 
    select(site, date_time, year, month, day, month_name, Temp_bot, Temp_mid, Temp_top)
  
  # return cleaned df ----
  return(temps_clean) # <3>
}
```
1.  create an input (aka **argument**) called `raw_data` inside `function()` (NOTE: you can name your argument however you'd like, but preferably something short and descriptive) 
2.  replace our hard-coded data frame name (e.g. `alegria` in the previous code chunk) in our cleaning pipeline with `raw_data`
3.  update the name of the object we save our clean data to (currently `alegria_clean`) to something a bit more generalized, like `temps_clean`, and `return()` our clean data frame object at the end of the function
4.  recall that part of our cleaning pipeline includes adding a column called `site`, with repeating values that are the site name; for now, let's just add some placeholder text ("_____") and we'll figure out how to make that text match up with the data in the next versions of our function

::: {.callout-note collapse="true"}
## What is `return()` and when is it necessary?

Oftentimes, we want a function to do some processing on whatever we provide it and then give us back the result. We use the `return()` function to do this in R. 

R **automatically** returns the the last output of a function -- here, it isn't necessary to `return(temps_clean)` since the `temps_clean` data frame is the last output of our `clean_ocean_temps()` function.  

An explicit `return()` is used to return a value immediately from a function. If it is *not* the last statement of a function, `return()` will prematurely end the function -- for example, if `x = 2`, the string, `"Positive"` will be returned (and the remaining `if else` statement will not be executed):

```{r}
#| eval: false
#| echo: true
#| message: false
#| warning: false
check_number <- function(x) {
  if (x > 0) {
    return("Positive")
  }
  else if (x < 0) {
    return("Negative")
  }
  else {
    return("Zero")
  }
}
```

I tend to include an explicit `return()` at the end of my functions because I think it makes it easier to read/understand the code, but check out [this interesting dialogue](https://stackoverflow.com/questions/11738823/explicitly-calling-return-in-a-function-or-not) on whether this is best practice or not.
:::

Lastly, let's make sure our function works. You can take a few approaches to trying out your work:

1.  **test in the console:** I do this a lot! It's fast and allows you to test out different things if you encounter any sticking points. You'll see me doing that while teaching
2.  **write and run your `functions_pipeline.R` script:** Begin writing your analysis script (here, that's `functions_pipeline.R`) -- below is one way you might consider setting up your script

Regardless, make sure to always rerun/re-source your function after making changes so that an updated version gets saved to your global environment!

```{r}
#| eval: false
#| echo: true
#| message: false
#| warning: false
# functions_pipeline.R #

#..........................load packages......................... 
library(readr) # <1>
library(dplyr) # <1>
library(tidyr) # <1>
library(chron) # <1>
library(naniar) # <1>

#........................source functions........................ 
source("utils/clean_ocean_temps.R") # <2>

#..........................import data........................... 
alegria <- read_csv(here::here("data", "raw_data", "alegria_mooring_ale_20210617.csv")) # <3>
mohawk <- read_csv(here::here("data", "raw_data", "mohawk_mooring_mko_20220330.csv")) # <3>
carpinteria <- read_csv(here::here("data", "raw_data", "carpinteria_mooring_car_20220330.csv")) # <3>

#...........................clean data........................... 
alegria_clean <- clean_ocean_temps(raw_data = alegria) # <4>
mohawk_clean <- clean_ocean_temps(raw_data = mohawk) # <4>
carpinteria_clean <- clean_ocean_temps(raw_data = carpinteria) # <4>
```
1. make sure any packages that your function relies on are installed/imported [^1] 
2. source your function into your script
3. read in your data
4. use your `clean_ocean_temps()` function to process your raw data  

[^1]: There are guidelines for handling dependencies (i.e. other packages that your function relies on) that are important to consider when building R packages. That's outside the scope of this workshop, but you can check out [R Packages (2e), Ch 11](https://r-pkgs.org/dependencies-mindset-background.html), by Hadley Wickham & Jenny Bryan, to learn more.

### **Version 2:** 

***The primary goal of `clean_ocean_temps() v2` is to create a `site` column that contains the correct site name.*** 

There are lots of creative ways to go about forming the site name that will get added as a repeating value to the `site` column, but the easiest and most explicit is likely adding a second function argument that takes the site name as a character string (e.g. `site_name = "alegria"`) 

Let's add that to our function:

```{r}
#| eval: false
#| echo: true
#| message: false
#| warning: false
clean_ocean_temps <- function(raw_data, site_name){ # <1>

  # format `site_name` ----
  site_name_formatted <- paste(str_to_title(site_name), "Reef") # <2>
    
  # clean data ----
  temps_clean <- raw_data |> 
    select(year, month, day, decimal_time, Temp_top, Temp_mid, Temp_bot) |>
    filter(year %in% c(2005:2020)) |> 
    mutate(site = rep(site_name_formatted)) |> # <3>
    unite(col = date, year, month, day, sep = "-", remove = FALSE) |> 
    mutate(time = times(decimal_time)) |> 
    unite(col = date_time, date, time, sep = " ", remove = TRUE) |> 
    mutate(date_time = as.POSIXct(date_time, "%Y-%m-%d %H:%M:%S", tz = "GMT"),
          year = as.factor(year),
          month = as.factor(month),
          day = as.numeric(day)) |>
    mutate(month_name = as.factor(month.name[month])) |>
    replace_with_na(replace = list(Temp_bot = 9999, Temp_top = 9999, Temp_mid = 9999)) |> 
    select(site, date_time, year, month, day, month_name, Temp_bot, Temp_mid, Temp_top)
  
  # return cleaned df ----
  return(temps_clean) 
  
}
```
1. use the `deparse()` and `substitute()` functions together to convert an object name (here, whichever data frame is passed in as `raw_data`) to a character string (`r fontawesome::fa("book-open-reader", fill = "#5A5A5A", a11y = "sem")` I referenced this [resource](https://statisticsglobe.com/convert-name-of-data-object-to-character-string-in-r )) 
2. we might specify in our function documentation that `site_name` takes the standard site name (as used by SBC LTER) as a character string, but what happens if a users uses different character casing than expected (e.g. `"alegria"`, `"MOHAWK"`, `"Carpinteria"`)? We can use a combination of `paste()` and the `{stringr}` package to format that value as we'd like it to appear in our `site` column -- here, that's converting it to Title Case and pasting `"Reef"` at the end (e.g. `"alegria"` will become `"Alegria Reef"`) ; `r fontawesome::fa("book-open-reader", fill = "#5A5A5A", a11y = "sem")` I referenced [this resource](https://stringr.tidyverse.org/reference/case.html)) 
3. substitute our formatted character string, `site_name_formatted`, in for the hard-coded site name in `mutate(site = rep("___ Reef"))`

Rerun the updated function and try using it to make sure the appropriate site name is added to the `site_name` column for each of the data sets:

```{r}
#| eval: false
#| echo: true
#| message: false
#| warning: false
# functions_pipeline.R #

#..........................load packages.........................  
library(readr)
library(dplyr) 
library(tidyr)
library(stringr) # <1>
library(chron)
library(naniar)

#........................source functions........................  
source("utils/clean_ocean_temps.R") 

#..........................import data...........................  
alegria <- read_csv(here::here("data", "raw_data", "alegria_mooring_ale_20210617.csv"))
mohawk <- read_csv(here::here("data", "raw_data", "mohawk_mooring_mko_20220330.csv"))
carpinteria <- read_csv(here::here("data", "raw_data", "carpinteria_mooring_car_20220330.csv"))

#...........................clean data...........................  
alegria_clean <- clean_ocean_temps(raw_data = alegria, site_name = "alegria") # <2>
mohawk_clean <- clean_ocean_temps(raw_data = mohawk, site_name = "MOHAWK") # <2>
carpinteria_clean <- clean_ocean_temps(raw_data = carpinteria, site_name = "Carpinteria") # <2>
```
1. remember to import the `{stringr}` package, since we'll be using `stringr::replace_with_na()` within our function
2. add in the `site_name` argument -- here, we demonstrate that how our function handles formatting, regardless of character case

### **Version 3:** 

***The primary goal of `clean_ocean_temps() v3 is to provide a way for users to select which temperature measurements (Temp_top, Temp_mid, Temp_bot) to include in the cleaned data frame.*** 

Our function works perfectly fine as-is, but let's say we don't always want all three temperature measurements (surface temperature (`Temp_top`), mid-column temperature (`Temp_mid`), and bottom temperature (`Temp_top`)) included in our cleaned data. We can build flexibility into our function by adding an argument that allows the user to select exactly which of the three temperature measurements to include in the resulting cleaned data frame. To do this, we'll make the following changes:

```{r}
#| eval: false
#| echo: true
#| message: false
#| warning: false
clean_ocean_temps <- function(raw_data, site_name, include_temps = c("Temp_top", "Temp_mid", "Temp_bot")){ # <1>
  
  # format `site_name` ----
  site_name_formatted <- paste(str_to_title(site_name), "Reef")
  
  # columns to select ---- 
  always_selected_cols <- c("year", "month", "day", "decimal_time") # <2>
  all_cols <- append(always_selected_cols, include_temps) # <3>
  
  # clean data ----
  temps_clean <- raw_data |> 
    select(all_of(all_cols)) |> # <4>
    filter(year %in% c(2005:2020)) |> 
    mutate(site = rep(site_name_formatted)) |> 
    unite(col = date, year, month, day, sep = "-", remove = FALSE) |> 
    mutate(time = times(decimal_time)) |> 
    unite(col = date_time, date, time, sep = " ", remove = TRUE) |> 
    mutate(date_time = as.POSIXct(date_time, "%Y-%m-%d %H:%M:%S", tz = "GMT"),
           year = as.factor(year),
           month = as.factor(month),
           day = as.numeric(day)) |>
    mutate(month_name = as.factor(month.name[month])) |>
    replace_with_na(replace = list(Temp_bot = 9999, Temp_top = 9999, Temp_mid = 9999)) |> 
    select(any_of(c("site", "date_time", "year", "month", "day", "month_name", "Temp_bot", "Temp_mid", "Temp_top"))) # <5>
  
  # return cleaned df ----
  return(temps_clean) 
}
```
1. add a third argument called `include_temps`, which defaults to including all three temperature variables (`Temp_top`, `Temp_mid`, `Temp_bot`)
2. create a vector of variable names that should *always* be selected
3. combine the "always selected" variables with the user-selected variables as specified using the `include_temps` argument
4. `select` columns based on variables names in our `all_cols` vector using `select(all_of())` (`r fontawesome::fa("book-open-reader", fill = "#5A5A5A", a11y = "sem")` I first got an error message [^2] when attempting to `select(all_cols)`, since an external vector alone can't be used to make selections; I then referenced this [resource](https://tidyselect.r-lib.org/reference/all_of.html)) 
5. make our last `select` call [^3], which is used to reorder columns, flexible enough to reorder temperature variables which may or may not be present using `select(any_of())` (`r fontawesome::fa("book-open-reader", fill = "#5A5A5A", a11y = "sem")` I again referenced this [resource](https://tidyselect.r-lib.org/reference/all_of.html))

[^2]: The error message I got when attempting to run `select(all_cols)`:

      `Using an external vector in selections was deprecated in tidyselect 1.1.0.`  
      `ℹ Please use 'all_of()' or 'any_of()' instead.`   
      `# Was:`   
      `data %>% select(all_cols)`  
      `# Now:`  
      `data %>% select(all_of(all_cols))`   
      `See <https://tidyselect.r-lib.org/reference/faq-external-vector.html>.`  
      
[^3]: `all_of()` and `any_of()` take a vector of character names, which is why all column names are in quotes, `""` (e.g. see `select(any_of(c("site", "date_time", "year", "month", "day", "month_name", "Temp_bot", "Temp_mid", "Temp_top"))))`

Rerun the updated function and try out our new `inlcude_temps` argument: 

```{r}
#| eval: false
#| echo: true
#| message: false
#| warning: false
# functions_pipeline.R #

#..........................load packages.........................  
library(readr)
library(dplyr) 
library(tidyr)
library(stringr) 
library(chron)
library(naniar)

#........................source functions........................ 
source("utils/clean_ocean_temps.R") 

#..........................import data...........................  
alegria <- read_csv(here::here("data", "raw_data", "alegria_mooring_ale_20210617.csv"))
mohawk <- read_csv(here::here("data", "raw_data", "mohawk_mooring_mko_20220330.csv"))
carpinteria <- read_csv(here::here("data", "raw_data", "carpinteria_mooring_car_20220330.csv"))

#...........................clean data........................... 
alegria_clean <- clean_ocean_temps(raw_data = alegria, site_name = "alegria", include_temps = c("Temp_bot")) # includes only `Temp_bot` # <1>
mohawk_clean <- clean_ocean_temps(raw_data = mohawk, site_name = "MOHAWK") # includes all three temp cols (`Temp_top`, `Temp_mid`, `Temp_bot`) by default # <1>
carpinteria_clean <- clean_ocean_temps(raw_data = carpinteria, site_name = "Carpinteria", include_temps = c("Temp_mid", "Temp_bot")) # includes only `Temp_mid` & `Temp_bot` # <1>
```
1. try using our new `include_temps` argument to return a cleaned data frame with a subset of the temperature measurement variables

::: {.callout-note}
## **Note:** `naniar::replace_with_na()` may throw a warning message 
However this will *not* halt execution -- if one or more of the temperature variables are missing (e.g. if we specify `include_temps = c("Temp_bot")`, you will get a warning that says, `Missing from data: Temp_top, Temp_mid`).
:::

### **Version 4:** 

***The primary goal of `clean_ocean_temps()` v4 is to build in checks to ensure that the data provided to the function is compatible with the cleaning pipeline (i.e. ensure that the correct columns are present.***

To wrap things up, we might consider adding an `if else` statement that checks to ensure that the data provided is suitable for our cleaning pipeline:

```{r}
#| eval: false
#| echo: true
#| message: false
#| warning: false
clean_ocean_temps <- function(raw_data, site_name, include_temps = c("Temp_top", "Temp_mid", "Temp_bot")){ 
  
  # if data contains these colnames, clean the script ----
  if(all(c("year", "month", "day", "decimal_time", "Temp_bot", "Temp_top", "Temp_mid") %in% colnames(raw_data))) { # <1>
    
    message("Cleaning data...") # <1>
    
    # format `site_name` ----
    site_name_formatted <- paste(str_to_title(site_name), "Reef")
    
    # columns to select ----
    standard_cols <- c("year", "month", "day", "decimal_time") 
    all_cols <- append(standard_cols, include_temps) 
    
    # clean data ----
    temps_clean <- raw_data |> 
      select(all_of(all_cols)) |> 
      filter(year %in% c(2005:2020)) |> 
      mutate(site = rep(site_name_formatted)) |> 
      unite(col = date, year, month, day, sep = "-", remove = FALSE) |> 
      mutate(time = times(decimal_time)) |> 
      unite(col = date_time, date, time, sep = " ", remove = TRUE) |> 
      mutate(date_time = as.POSIXct(date_time, "%Y-%m-%d %H:%M:%S", tz = "GMT"),
             year = as.factor(year),
             month = as.factor(month),
             day = as.numeric(day)) |>
      mutate(month_name = as.factor(month.name[month])) |>
      replace_with_na(replace = list(Temp_bot = 9999, Temp_top = 9999, Temp_mid = 9999)) |> 
      select(any_of(c("site", "date_time", "year", "month", "day", "month_name", "Temp_bot", "Temp_mid", "Temp_top")))
    
    # return cleaned df ----
    return(temps_clean) # <1>
    
  } else { # <2>
    
    stop("The data frame provided does not include the necessary columns. Double check your data!") # <2>
    
  } # <2>
  
}
```
1. add an `if else` statement that checks whether the necessary columns are present in the raw data -- if yes, proceed with data cleaning
2. if no, throw an error message

Let's rerun our function and try it out one last time: 

```{r}
#| eval: false
#| echo: true
#| message: false
#| warning: false
# functions_pipeline.R #

#..........................load packages.........................  
library(readr)
library(dplyr) 
library(tidyr)
library(stringr) 
library(chron)
library(naniar)

#........................source functions........................ 
source("utils/clean_ocean_temps.R") 

#..........................import data...........................  
alegria <- read_csv(here::here("data", "raw_data", "alegria_mooring_ale_20210617.csv"))
mohawk <- read_csv(here::here("data", "raw_data", "mohawk_mooring_mko_20220330.csv"))
carpinteria <- read_csv(here::here("data", "raw_data", "carpinteria_mooring_car_20220330.csv"))

#...........................clean data...........................  
alegria_clean <- clean_ocean_temps(raw_data = alegria, site_name = "alegria", include_temps = c("Temp_bot")) # <1>
mohawk_clean <- clean_ocean_temps(raw_data = mohawk, site_name = "MOHAWK") # <1>
carpinteria_clean <- clean_ocean_temps(raw_data = carpinteria, site_name = "Carpinteria", include_temps = c("Temp_mid", "Temp_bot")) # <1>
penguins_clean <- clean_ocean_temps(raw_data = palmerpenguins::penguins) # <2>
```
1. these three should work as intended
2. this one should throw an error!

## **ii:** Write a function to **plot** data sets

Similar to what we did for our cleaning code, let's first identify which parts of the plotting code need to be generalized/made "flexible" so that any of our three cleaned data frames can be passed to it for plotting. For us, that's the **name of the clean data frame** and the **site name** that appears in the plot title (see annotation notes below the rendered code):

```{r}
#| eval: false
#| echo: true
#| message: false
#| warning: false
alegria_plot <- alegria_clean |> # <1>
  group_by(month_name) |> 
  ggplot(aes(x = Temp_bot, y = month_name, fill = after_stat(x))) +
  geom_density_ridges_gradient(rel_min_height = 0.01, scale = 3) + 
  scale_x_continuous(breaks = c(9, 12, 15, 18, 21)) +
  scale_y_discrete(limits = rev(month.name)) + 
  scale_fill_gradientn(colors = c("#2C5374","#778798", "#ADD8E6", "#EF8080", "#8B3A3A"), name = "Temp. (°C)") +
  labs(x = "Bottom Temperature (°C)",
       title = "Bottom Temperatures at Alegria Reef, Santa Barbara, CA", # <2>
       subtitle = "Temperatures (°C) aggregated by month from 2005 - 2020") +
  ggridges::theme_ridges(font_size = 13, grid = TRUE) +
  theme(
    axis.title.y = element_blank()
  )

alegria_plot
```
1. the name of the clean data frame needs to be generalized
2. the site name that appears in the plot title needs to be generalized

### **Version 1:** 

***The primary goal of `plot_ocean_temps` v1 is to get a basic plotting function working.***

Now we can begin to building our function. Let's again start by creating the skeleton of our function, which we’ll call `plot_ocean_temps()`:

```{r}
#| eval: false
#| echo: true
#| message: false
#| warning: false
plot_ocean_temps <- function(){
  
}
```

...then copy and paste our plotting code for the clean Alegria Reef data into the body of the function (i.e. within the curly brackets, `{}`).

```{r}
#| eval: false
#| echo: true
#| message: false
#| warning: false
plot_ocean_temps <- function() {

  alegria_plot <- alegria_clean |> 
    group_by(month_name) |> 
    ggplot(aes(x = Temp_bot, y = month_name, fill = after_stat(x))) +
    geom_density_ridges_gradient(rel_min_height = 0.01, scale = 3) + 
    scale_x_continuous(breaks = c(9, 12, 15, 18, 21)) +
    scale_y_discrete(limits = rev(month.name)) + 
    scale_fill_gradientn(colors = c("#2C5374","#778798", "#ADD8E6", "#EF8080", "#8B3A3A"), name = "Temp. (°C)") +
    labs(x = "Bottom Temperature (°C)",
         title = "Bottom Temperatures at Alegria Reef, Santa Barbara, CA",
         subtitle = "Temperatures (°C) aggregated by month from 2005 - 2020") +
    ggridges::theme_ridges(font_size = 13, grid = TRUE) +
    theme(
      axis.title.y = element_blank()
    )
  
}
```

Just like in `clean_ocean_temps()`, we want the ability to provide our function with any three of our cleaned data sets for plotting. To do this, we'll make the following modifications:

```{r}
#| eval: false
#| echo: true
#| message: false
#| warning: false
plot_ocean_temps <- function(clean_data) { # <1>

  # plot data ----
  temps_plot <- clean_data |> # <2>
    group_by(month_name) |> 
    ggplot(aes(x = Temp_bot, y = month_name, fill = after_stat(x))) +
    geom_density_ridges_gradient(rel_min_height = 0.01, scale = 3) + 
    scale_x_continuous(breaks = c(9, 12, 15, 18, 21)) +
    scale_y_discrete(limits = rev(month.name)) + 
    scale_fill_gradientn(colors = c("#2C5374","#778798", "#ADD8E6", "#EF8080", "#8B3A3A"), name = "Temp. (°C)") +
    labs(x = "Bottom Temperature (°C)",
         title = "Bottom Temperatures at _____, Santa Barbara, CA", # <4>
         subtitle = "Temperatures (°C) aggregated by month from 2005 - 2020") +
    ggridges::theme_ridges(font_size = 13, grid = TRUE) +
    theme(
      axis.title.y = element_blank()
    )
  
  # return plot ----
  return(temps_plot) # <3>
  
}
```
1. create an input (aka argument) called `clean_data` inside `function()` (NOTE: you can name your argument however you’d like, but preferably something short and descriptive)
2. replace our hard-coded data frame name (e.g. `alegria_clean` in the code above) in our plotting pipeline with our new argument, `clean_data`
3. update the name of the object we save our plot output to (currently `alegria_plot`) to something a bit more generalized, like `temps_plot` and `return()` our plot object at the end
4. and add some placeholder text ("_____") in the `title` field where the site name should be -- we'll make this flexible in our next version

Now let's make sure our function works. Run the function so that it's saved to our global environment, then use it to try plotting our cleaned data sets:

```{r}
#| eval: false
#| echo: true
#| message: false
#| warning: false
# functions_pipeline.R #

#..........................load packages.........................  
library(readr)
library(dplyr) 
library(tidyr)
library(stringr) 
library(chron)
library(naniar)
library(ggplot2) # <1>
library(ggridges) # <1>

#........................source functions........................  
source("utils/clean_ocean_temps.R") 
source("utils/plot_ocean_temps.R")

#..........................import data...........................  
alegria <- read_csv(here::here("data", "raw_data", "alegria_mooring_ale_20210617.csv"))
mohawk <- read_csv(here::here("data", "raw_data", "mohawk_mooring_mko_20220330.csv"))
carpinteria <- read_csv(here::here("data", "raw_data", "carpinteria_mooring_car_20220330.csv"))

#...........................clean data...........................  
alegria_clean <- clean_ocean_temps(raw_data = alegria, site_name = "alegria", include_temps = c("Temp_bot")) 
mohawk_clean <- clean_ocean_temps(raw_data = mohawk, site_name = "MOHAWK") 
carpinteria_clean <- clean_ocean_temps(raw_data = carpinteria, site_name = "Carpinteria", include_temps = c("Temp_mid", "Temp_bot")) 

#............................plot data...........................
alegria_plot <- plot_ocean_temps(clean_data = alegria_clean) # <2>
mohawk_plot <- plot_ocean_temps(clean_data = mohawk_clean) # <2>
carpinteria_plot <- plot_ocean_temps(clean_data = carpinteria_clean) # <2>
```
1. make sure any additional packages that our `plot_ocean_temps()` function relies on are installed/imported 
2. **Reminder:** `plot_ocean_temps()` is written so that it plots **bottom temperatures (`Temp_bot`)**, so make sure that your cleaned data sets have this variable present (i.e. make sure that `include_temps` is set to it's default *or* that `Temp_bot` is explicitly included)

### **Version 2:**

***The primary goal of `plot_ocean_temps() v2` is to to create a plot `title` that contains the correct side name.*** 

In this next version, let's figure out how to update the plot `title` so that the appropriate site name appears each time a different data set is plotted. Luckily for us, we've already added a `site` column to our `*_clean` data, which contains the site name as a character string. We can extract this site name by adding the following:

```{r}
#| eval: false
#| echo: true
#| message: false
#| warning: false
plot_ocean_temps <- function(clean_data) { 

  # get site name for plot title ----
  site_name <- unique(clean_data$site) # <1>
  
  # plot data ----
  temp_plot <- clean_data |> 
    group_by(month_name) |> 
    ggplot(aes(x = Temp_bot, y = month_name, fill = after_stat(x))) +
    ggridges::geom_density_ridges_gradient(rel_min_height = 0.01, scale = 3) + 
    scale_x_continuous(breaks = c(9, 12, 15, 18, 21)) +
    scale_y_discrete(limits = rev(month.name)) + 
    scale_fill_gradientn(colors = c("#2C5374","#778798", "#ADD8E6", "#EF8080", "#8B3A3A"), name = "Temp. (°C)") +
    labs(x = "Bottom Temperature (°C)",
         title = paste("Bottom Temperatures at ",  site_name, ", Santa Barbara, CA", sep = ""), # <2>
         subtitle = "Temperatures (°C) aggregated by month from 2005 - 2022") +
    ggridges::theme_ridges(font_size = 13, grid = TRUE) +
    theme(
      axis.title.y = element_blank()
    )
  
  # return plot ----
  return(temp_plot)
}
```
1. use the `unique()` function to get the site name for a given data set; recall that we should only have 1 unique site name per data set
2. use the `paste()` function to construct a `title` using our extracted site name

Rerun and try out your function again to make sure it works as expected:

```{r}
#| eval: false
#| echo: true
#| message: false
#| warning: false
# functions_pipeline.R #

#..........................load packages.........................  
library(readr)
library(dplyr) 
library(tidyr)
library(stringr) 
library(chron)
library(naniar)
library(ggplot2)
library(ggridges) 

#........................source functions........................  
source("utils/clean_ocean_temps.R") 
source("utils/plot_ocean_temps.R") # <1>

#..........................import data...........................  
alegria <- read_csv(here::here("data", "raw_data", "alegria_mooring_ale_20210617.csv"))
mohawk <- read_csv(here::here("data", "raw_data", "mohawk_mooring_mko_20220330.csv"))
carpinteria <- read_csv(here::here("data", "raw_data", "carpinteria_mooring_car_20220330.csv"))

#...........................clean data...........................  
alegria_clean <- clean_ocean_temps(raw_data = alegria, site_name = "alegria", include_temps = c("Temp_bot")) 
mohawk_clean <- clean_ocean_temps(raw_data = mohawk, site_name = "MOHAWK") 
carpinteria_clean <- clean_ocean_temps(raw_data = carpinteria, site_name = "Carpinteria", include_temps = c("Temp_mid", "Temp_bot")) 

#............................plot data...........................
alegria_plot <- plot_ocean_temps(clean_data = alegria_clean) 
mohawk_plot <- plot_ocean_temps(clean_data = mohawk_clean) 
carpinteria_plot <- plot_ocean_temps(clean_data = carpinteria_clean) 
```
1. **remember** to re-load your function after you make changes!

## **iii:** Putting it all together

Okay, now let's bring all these pieces together! Our revised scripts might look something like this:

::: {.panel-tabset}

## `functions_pipeline.R`

```{r}
#| eval: false
#| echo: true
#| message: false
#| warning: false
#..........................load packages......................... 
library(readr)
library(dplyr) 
library(tidyr)
library(stringr) 
library(chron)
library(naniar)
library(ggplot2)
library(ggridges) 

#........................source functions........................  
source("utils/clean_ocean_temps.R") 
source("utils/plot_ocean_temps.R") 

#..........................import data...........................  
alegria <- read_csv(here::here("data", "raw_data", "alegria_mooring_ale_20210617.csv"))
mohawk <- read_csv(here::here("data", "raw_data", "mohawk_mooring_mko_20220330.csv"))
carpinteria <- read_csv(here::here("data", "raw_data", "carpinteria_mooring_car_20220330.csv"))

#...........................clean data...........................  
alegria_clean <- clean_ocean_temps(raw_data = alegria, site_name = "alegria", include_temps = c("Temp_bot")) 
mohawk_clean <- clean_ocean_temps(raw_data = mohawk, site_name = "MOHAWK") 
carpinteria_clean <- clean_ocean_temps(raw_data = carpinteria, site_name = "Carpinteria", include_temps = c("Temp_mid", "Temp_bot"))

#............................plot data...........................
alegria_plot <- plot_ocean_temps(clean_data = alegria_clean) 
mohawk_plot <- plot_ocean_temps(clean_data = mohawk_clean) 
carpinteria_plot <- plot_ocean_temps(clean_data = carpinteria_clean) 
```

## `utils/clean_ocean_temps.R`

::: {.callout-note}
## Use a roxygen skeleton to document your function
Even if you don't plan to publish your function as part of a package, documenting your work is still a critical part of reproducibility and usability. This may be done in more informal ways, such as code annotations and text explanations in RMarkdown documents, for example. You may also consider more formal documentation -- the [`{roxygen2}`](https://cran.r-project.org/web/packages/roxygen2/vignettes/roxygen2.html) package helps to make that process easier. Click anywhere inside your function, then choose Code > Insert Roxygen Skeleton to get started. 
:::

```{r}
#| eval: false
#| echo: true
#| message: false
#| warning: false
#' Process CTD/ADCP temperature data
#'
#' @param raw_data data frame of CTD/ADCP data collected at SBC LTER site moorings; search for data on the EDI Data Portal (http://portal.edirepository.org:80/nis/simpleSearch?defType=edismax&q=SBC+LTER%5C%3A+Ocean%5C%3A+Currents+and+Biogeochemistry%5C%3A+Moored+CTD+and+ADCP+data&fq=-scope:ecotrends&fq=-scope:lter-landsat*&fl=id,packageid,title,author,organization,pubdate,coordinates&debug=false)
#' @param site_name Santa Barbara Coastal LTER site name as a character string
#' @param include_temps vector of character strings that includes one or more of the following variable names: Temp_top, Temp_mid, Temp_top
#'
#' @return a data frame
#' @export
#'
#' @examples
#' my_clean_df <- clean_ocean_temps(raw_data = my_raw_df, include_temps = c("Temp_bot"))
clean_ocean_temps <- function(raw_data, site_name, include_temps = c("Temp_top", "Temp_mid", "Temp_bot")){ 
  
  # if data contains these colnames, clean the script ----
  if(all(c("year", "month", "day", "decimal_time", "Temp_bot", "Temp_top", "Temp_mid") %in% colnames(raw_data))) { 
    
    message("Cleaning data...") 
    
    # format site name ----
    site_name_formatted <- paste(str_to_title(site_name), "Reef")
    
    # columns to select ----
    standard_cols <- c("year", "month", "day", "decimal_time") 
    all_cols <- append(standard_cols, include_temps) 
    
    # clean data ----
    temps_clean <- raw_data |> 
      select(all_of(all_cols)) |> 
      filter(year %in% c(2005:2020)) |> 
      mutate(site = rep(site_name_formatted)) |> 
      unite(col = date, year, month, day, sep = "-", remove = FALSE) |> 
      mutate(time = times(decimal_time)) |> 
      unite(col = date_time, date, time, sep = " ", remove = TRUE) |> 
      mutate(date_time = as.POSIXct(date_time, "%Y-%m-%d %H:%M:%S", tz = "GMT"),
             year = as.factor(year),
             month = as.factor(month),
             day = as.numeric(day)) |>
      mutate(month_name = as.factor(month.name[month])) |>
      replace_with_na(replace = list(Temp_bot = 9999, Temp_top = 9999, Temp_mid = 9999)) |> 
      select(any_of(c("site", "date_time", "year", "month", "day", "month_name", "Temp_bot", "Temp_mid", "Temp_top")))
    
    # return cleaned df ----
    return(temps_clean) 
    
  } else { 
    
    stop("The data frame provided does not include the necessary columns. Double check your data!") 
    
  }
  
}
```

## `utils/plot_ocean_temps.R`

::: {.callout-note}
## Use a roxygen skeleton to document your function
Even if you don't plan to publish your function as part of a package, documenting your work is still a critical part of reproducibility and usability. This may be done in more informal ways, such as code annotations and text explanations in RMarkdown documents, for example. You may also consider more formal documentation -- the [`{roxygen2}`](https://cran.r-project.org/web/packages/roxygen2/vignettes/roxygen2.html) package helps to make that process easier. Click anywhere inside your function, then choose Code > Insert Roxygen Skeleton to get started. 
:::

```{r}
#| eval: false
#| echo: true
#| message: false
#| warning: false
#' Create ridge line plot of CTD/ADCP bottom temperature data
#'
#' @param clean_data a data frame that has been pre-processed using clean_ocean_temp()
#'
#' @return a plot object
#' @export
#'
#' @examples
#' my_plot <- plot_ocean_temps(clean_data = my_clean_df)
plot_ocean_temps <- function(clean_data) { 
  
  # get site name for plot title ----
  site_name <- unique(clean_data$site)
  
  # plot data ----
  temp_plot <- clean_data |> 
    group_by(month_name) |> 
    ggplot(aes(x = Temp_bot, y = month_name, fill = after_stat(x))) +
    ggridges::geom_density_ridges_gradient(rel_min_height = 0.01, scale = 3) + 
    scale_x_continuous(breaks = c(9, 12, 15, 18, 21)) +
    scale_y_discrete(limits = rev(month.name)) + 
    scale_fill_gradientn(colors = c("#2C5374","#778798", "#ADD8E6", "#EF8080", "#8B3A3A"), name = "Temp. (°C)") +
    labs(x = "Bottom Temperature (°C)",
         title = paste("Bottom Temperatures at ", site_name, ", Santa Barbara, CA", sep = ""), 
         subtitle = "Temperatures (°C) aggregated by month from 2005 - 2022") +
    ggridges::theme_ridges(font_size = 13, grid = TRUE) +
    theme(
      axis.title.y = element_blank()
    )
  
  # return plot ----
  return(temp_plot)
}
```

:::